---
title: "Penguin Assignment Q4"
author: "Student No. 1336616"
date: "04/12/2022"
output: 
  html_document:
    theme: united
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment="", prompt=TRUE)
```

# Set working directory 
```{r SetWorkingDirectory, eval = FALSE, include = TRUE, warning = FALSE}
setwd(yourpathhere) # set your working directory 
getwd() # check working directory 
```

# Load in the functions and data

Load in the functions

```{r LoadFunctions}
source("functions/libraries.r") # libraries to load
source("functions/cleaning.r") # functions for cleaning and sub-setting the data
source("functions/plotting.r") # functions for plotting and saving the data
```

Load in the penguins data

```{r, PreservingRawData}
write.csv(penguins_raw, paste0(("data-raw/penguins_raw.csv"))) # this is the code I used to write the CSV
```

```{r LoadData}
penguins_raw <- read.csv("data-raw/penguins_raw.csv") # reading in the data
```

# Clean the penguins data

```{r CleanData}
penguins_clean <- cleaning(penguins_raw) # cleaning the data 

# I used the following code to save the "clean data" CSV file 
date <- Sys.Date() # getting the date for saving 
write.csv(penguins_clean, paste("data-clean/", date, "_penguins_clean.csv",sep="")) # preserving the clean data with a date
```

```{r SubsetData}
penguins_culmen <- remove_empty_culmen_depth(penguins_clean) # sub-setting data to include Gentoo penguin species that are not NA for the culmen depth
View(penguins_culmen) # visually checking the data 
```

# Model and statistical test

We are modelling a linear regression to analyse the relationship between body mass (g) and  culmen depth (mm) for Gentoo penguins. 

```{r FittingRegressionModel, message = FALSE}
culmen_mod1 <- lm(culmen_depth_mm ~ body_mass_g, penguins_culmen) # fitting the linear regression model 
```

```{r DiagnosticPlots, figures-side, fig.show="hold", out.width="50%", message=FALSE, results = "hide"}
diagnosticplots <- plot(culmen_mod1) # checking that the model fits our assumptions of heterogeneity by looking at the diagnostic plots. This is so that we can run the ANCOVA.
diagnosticplots
```


The diagnostic plots appear to fit our assumptions for ANOVA well (normal distribution and residual heterogeneity). As such, we can continue with the ANOVA test.

```{r SummaryTable}
summary(culmen_mod1) # getting a summary table for the model
```

Basic analysis and discussion from the summary table:

* The equation for the graph is y = 0.0014x + 7.9. It is a positive linear correlation 
* Multiple R^2^ is 0.52, and adjusted R^2^ is 0.52. This means that culmen depth explains about 50% of the variance in body mass for the Gentoo penguins
* This finding is statistically significant, with a P-value of < 2.2x10^16^
* Evidence to reject the null hypothesis, that the slope in the species data between body mass and culmen depth is not equal to 0

# Saving the linear regression summary
```{r, results='hide', message = FALSE}
sink(file = "other-outputs/linear-regression-summary.txt")
summary(culmen_mod1)
sink()
```

```{r ANOVATable}
anova(culmen_mod1) # running an anova for the model
```

Basic analysis and discussion from the ANOVA table:

* P-value is less than 0.05 (< 2.2x10^16^)
* We can use body mass to predict culmen depth - as body mass increases, so does culmen depth. This is evidence to reject the null hypothesis that one does not significantly predict the other
* Overall indicates that Gentoo penguins with larger body mass will have a greater culmen depth, which may be down to their basic biology (i.e. larger penguins have larger features due to the fact that it is a correlated phenotype)

# Saving the ANOVA summary
```{r, results='hide', message = FALSE}
sink(file = "other-outputs/ANOVA-summary.txt")
anova(culmen_mod1)
sink()
```
# Creating a figure for the model

```{r CreateFigure, fig.crop = FALSE}
culmen_scatterplot <- plot_culmen_figure(penguins_culmen) # using the plotting function
culmen_scatterplot # viewing the scatterplot
```

# Saving the figures
```{r SaveFigurePNG, message = FALSE, results = "hide"}
save_culmen_plot_png(penguins_culmen, "figures/fig01_culmen_report.png", 
                    size = 20, res = 600, scaling = 1.3) # using the saving functions to save figure as png, tailored to a report size, resolution, and scaling 

```

```{r SaveFigureSVG, message = FALSE, results = "hide"}
save_culmen_plot_svg(penguins_culmen, "figures/fig01_culmen_vector.svg", 
                      size = 20, scaling = 1.3) # using the saving functions to save the figure as an svg
```

# Saving the software versions used
```{r message = FALSE, results = "hide"}
sink(file = "other-outputs/package-versions.txt")
sessionInfo()
sink() # saving the software used
```
